### 2015-02-27 icfp15-servo/abstract.tex

%!TEX root = paper.tex
%
\begin{abstract}
All modern web browsers --- Internet Explorer, Firefox, Chrome, Opera, and Webkit ---
have a core rendering engine written in \Cplusplus.
This language choice was made because it affords the systems programmer complete control
of the underlying hardware features and memory in use plus provides a transparent
compilation model.

>>>>

%!TEX root = paper.tex
%
\begin{abstract}
All modern web browsers --- Internet Explorer, Firefox, Chrome, Opera, and Webkit ---
have a core rendering engine written in \Cplusplus.
This language choice was made because it affords the systems programmer complete control
of the underlying hardware features and memory in use and it provides a transparent
compilation model.

### 2015-02-27 icfp15-servo/intro.tex

Servo~\cite{SERVO} is a new web browser engine designed to address the major environment and 
architectural changes over the last decade.
The goal of the Servo project is to produce a browser that enables new applications to be authored
against the web platform that run with more safety, better performance, and better power usage
than in current browsers.
To address memory-related safety issues, we are using a new systems programming language,
Rust~\cite{RUST}.
For parallelism and power, we scale across a wide variety of hardware by building either data-
or task-parallelism, as appropriate, to each of the parts of the web platform.
Additionally, we are improving concurrency by reducing the simultaneous access to data
structures and using a message-passing architecture between components such as the
JavaScript engine and the rendering engine that is painting graphics to the screen.
Servo is currently over 300k lines of code and implements enough of the web to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser and its associated libraries.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of programming languages and runtime research.
In this experience report, we discuss the design and architecture of a modern web 
browser engine, show how modern programming language techniques --- many of which
originated in the functional programming community --- address these design 
constraints, and also touch on ongoing challenges and areas of research where we
would welcome additional community input.

>>>>

Servo~\cite{SERVO} is a new web browser engine designed to address the major environment and 
architectural changes over the last decade.
The goal of the Servo project is to produce a browser that enables new applications to be authored
against the web platform that run with more safety, better performance, and better power usage
than in current browsers.
To address memory-related safety issues, we are using a new systems programming language,
Rust~\cite{RUST}.
For parallelism and power, we scale across a wide variety of hardware by building either data-
or task-parallelism, as appropriate, into each of the parts of the web platform.
Additionally, we are improving concurrency by reducing the simultaneous access to data
structures and using a message-passing architecture between components such as the
JavaScript engine and the rendering engine that is painting graphics to the screen.
Servo is currently over 300k lines of code and implements enough of the web to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser and its associated libraries.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of programming languages and runtime research.
In this experience report, we discuss the design and architecture of a modern web 
browser engine, show how modern programming language techniques --- many of which
originated in the functional programming community --- address these design 
constraints, and also touch on ongoing challenges and areas of research where we
would welcome additional community input.

### 2015-02-27 icfp15-servo/rel.tex

%!TEX root = paper.tex
%
\section{Related browser research}
\label{sec:rel}
The ZOOMM browser was an effort at Qualcomm Research to build a parallel browser, also
focused on multicore mobile devices~\cite{ZOOMM}.
This browser includes many features that we have not yet implemented in Servo, particularly
around resource fetching.
They also wrote their own JavaScript engine, which we have not done.
Servo and ZOOMM share an extremely concurrent architecture --- both have script, layout,
rendering, and the user interface operating concurrently from one another in order to maximize
interactivity for the user.
Parallel layout in Servo is one major area that was not investigated in the ZOOM browser.
The other major difference is that Servo is implemented in Rust, whereas ZOOM is written
in \Cplusplus, similarly to most modern browser engines.

Ras Bodik's group at the University of California Berkely worked on a parallel browsing
project (funded in part by Mozilla Research) that focused on improving the parallelism
of layout~\cite{parallel-layout}.
Instead of our approach to parallel layout, which focuses on multiple parallel tree
traversals, they modeled a subset of CSS using attribute grammars.
They showed significant speedups with their system over a reimplementation of Safari's
algorithms, but we have not used this approach due to questions of whether it is possible
to use attribute grammars to both accurately model the web as it is implemented today
and to support new features as they are added.
Servo uses a very similar CSS selector matching algorithm to theirs.

>>>>

%!TEX root = paper.tex
%
\section{Related browser research}
\label{sec:rel}
The ZOOMM browser was an effort at Qualcomm Research to build a parallel browser, also
focused on multicore mobile devices~\cite{ZOOMM}.
This browser includes many features that we have not yet implemented in Servo, particularly
around resource fetching.
They also wrote their own JavaScript engine, which we have not done.
Servo and ZOOMM share an extremely concurrent architecture --- both have script, layout,
rendering, and the user interface operating concurrently from one another in order to maximize
interactivity for the user.
Parallel layout in Servo is one major area that was not investigated in the ZOOMM browser.
The other major difference is that Servo is implemented in Rust, whereas ZOOMM is written
in \Cplusplus, similarly to most modern browser engines.

Ras Bodik's group at the University of California Berkeley worked on a parallel browsing
project (funded in part by Mozilla Research) that focused on improving the parallelism
of layout~\cite{parallel-layout}.
Instead of our approach to parallel layout, which focuses on multiple parallel tree
traversals, they modeled a subset of CSS using attribute grammars.
They showed significant speedups with their system over a reimplementation of Safari's
algorithms, but we have not used this approach due to questions of whether it is possible
to use attribute grammars to both accurately model the web as it is implemented today
and to support new features as they are added.
Servo uses a very similar CSS selector matching algorithm to theirs.
